---
title: "World Development Indicators (WDI) Health Dataset"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(rworldmap)
library(ggthemes)
library(corrplot)
set.seed(5346)
setwd("C:/Users/xmint/Desktop/Health_R")
```

### --- Health ---

This set of visualisations is based on World Bank's world development indicators (WDI) released for the year 2020. 

Variables include:

  * Country Name and Code 
  * Life expectancy at birth, total (years)
  * Life expectancy at birth, female (years)
  * Life expectancy at birth, male (years)
  * Mortality rate, infant (per 1000 live births)
  * Urban population (% of total population)
  * Current Health Expenditure (% of GDP)

```{r echo=FALSE}
### Load data file ###
# Download CSV file from https://qog.pol.gu.se/data/datadownloads/qogstandarddata
qog <- read.csv("qog_std_cs_jan20.csv",header=TRUE)
qog <- qog %>% select(ccode,cname,ccodealp,ccodecow,wdi_lifexp,wdi_lifexpf, wdi_lifexpm, wdi_mortinf,wdi_popurb,wdi_chexppgdp)
```

### 1 Life expectancy distribution by Gender

```{r echo=FALSE}
### Life expectancy distribution for female and male
life <- rbind(data.frame(life_expectancy = qog$wdi_lifexpf,sex="F"),data.frame(life_expectancy = qog$wdi_lifexpm,sex="M")) %>% na.omit()
ggplot(life, aes(life_expectancy)) + geom_histogram(aes(y=..density..,color = sex, fill = sex), alpha = 0.4, position = "identity",binwidth = 1) +
  scale_fill_manual(values=c("#00AFBB", "#E7B800")) + scale_color_manual(values=c("#00AFBB", "#E7B800")) + geom_density(aes(life_expectancy,color = sex), size = 1) + 
  xlab("Life Expectancy") + ylab("Density")
```

The life expectancy of males around the world is generally lower than the life expectancy of females. 

### 2 Life expectancy (World Map)
```{r echo=FALSE,message=FALSE,warning=FALSE,error=FALSE}
### Life expectancy distribution for female and male
life_country <- data.frame(region = qog$ccodealp, life_expectancy = qog$wdi_lifexp)

# World map 
mapped_data <- joinCountryData2Map(life_country, joinCode = "ISO3", nameJoinColumn = "region")

# Trying to use ggplot instead to plot a nicer graph
new_mapping <- data.frame(region = mapped_data$NAME,life_expectancy = mapped_data$life_expectancy)
wmap <- map_data("world")
world_base <- ggplot(data = wmap, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + geom_polygon(color = "black", fill = "gray", lty=.1)
#Filling up some value missing for big countries 
map_life <- left_join(wmap, new_mapping, by = "region")
map_life$life_expectancy[map_life$region == "Czech Republic"] <- 78.824387
map_life$life_expectancy[map_life$region == "UK"] <- 81.304878
map_life$life_expectancy[map_life$region == "South Korea"] <- 82.155853
map_life$life_expectancy[map_life$region == "North Korea"] <- 70.952072
map_life$life_expectancy[map_life$region == "USA"] <- 78.741463
map_life$life_expectancy[map_life$region == "Democratic Republic of the Congo"] <- 58.748611
map_life$life_expectancy[map_life$region == "Republic of Congo"] <- 63.515316
map_life$life_expectancy[map_life$region == "Central African Republic"] <- 50.591049
map_life$life_expectancy[map_life$region == "South Sudan"] <- 55.796486

#Final World Map graph
world_base + 
  geom_polygon(data = map_life, aes(fill = life_expectancy), color = "white", lty=.1) + 
  geom_polygon(color = "white", fill = NA, lty=.1) + 
  scale_fill_gradient(name = "Life Expectancy", low="#d0587e", high="#2887a1", na.value = "grey50") + 
  theme(legend.position='top') + theme_fivethirtyeight()
  
```

Continents such as Europe, North America and Australia have higher life expectancy of more than 75 (shown by the solid blue) as compared to Africa which has lower life expectancy of 60 and below (shown by the solid red). 

### 3 Correlation plot between selected variables

```{r echo=FALSE}

group_urban <- data.frame(life_expectancy = qog$wdi_lifexp, mortality_rate = qog$wdi_mortinf, urban_percentage = qog$wdi_popurb, health_expenditure = qog$wdi_chexppgdp) %>% na.omit
group_urban$urban50 <- NA
group_urban$urban50[group_urban$urban_percentage < 50] = "rural"
group_urban$urban50[group_urban$urban_percentage >= 50] = "urban"

# Correlation plot 
actual_frame <- select(group_urban,-urban50) %>% na.omit()
cor_frame <- cor(actual_frame)

corrplot(cor_frame,method = "number", tl.cex=1, tl.srt = 45, tl.col = "black", order="hclust", type ="upper" )
```

Life expectancy and mortality rate have a strong negative correlation of -0.93. We can also see that percentage of urban population has a stronger correlation with mortality rate and life expectancy as compared to health expenditure of the country. Percentage of urban population has a moderate negative correlation with mortality rate and a moderate positive correlation with life expectancy.

### 4 Relationship between mortality rate, life expectancy and percentage of urban population.
```{r echo=FALSE}

### Mortality Rate against Life expectancy (with Urban Percentage) 

# To see some regression outputs
# y <- lm(wdi_lifexp ~ wdi_mortinf + wdi_popurb, qog)
# summary(y)
# x <- lm(wdi_lifexp ~ wdi_mortinf + wdi_popurb + wdi_chexppgdp, qog)
# summary(x)

# Graph plot
ggplot(group_urban,aes(y=life_expectancy,x=mortality_rate)) + geom_point(aes(color = urban_percentage)) + 
  scale_color_gradient(low = "blue",high = "red") + geom_smooth(method = lm) +
  xlab("Mortality Rate") + ylab("Life Expectancy")

```

A linear model with 95% confidence interval (blue line) is fitted onto the scatter plot. We see that countries with higher life expectancy tends to have lower mortality rate, perhaps due to better healthcare. In addition, countries with high life expectancy and low mortality rate (top left) usually also have a larger percentage of urban population (shown by the brighter red circles). This reinforces the insights from the correlation plot.
